The Galactic Midplane is Not a Plane: Implications for dynamical analysis with {\em Gaia} data and beyond

### 2019-04-22 paper/ms.tex
Add velocities to orbit
Add a little bit about constancy of frequencies

Under this approximation, we can write down the vertical and radial components
of the orbits as,
\beq\label{eq:orbits_epi}
\begin{split}
z(t) &= A_z \sin{(\kappa t + \alpha)}\\
R(t) &= A_R \sin{(\nu t + \beta)}\text{,}
\end{split}
\eeq
where $\kappa$ and $\nu$ are the vertical and radial (or epicyclic)
frequencies, $A_z$ and $A_R$ are the amplitudes of the stellar motion in the
vertical and radial coordinates, and $\alpha$ and $\beta$ are arbitrary
constants. We can write down the actions in this approximation as
\citep[Section~3.5.3b][]{2008gady.book.....B},
\beq\label{eq:actions_epi}
\begin{split}
L_z &= R_g v_c \\
J_r &= \kappa A_R^2 \\
J_z &= \nu A_z^2\text{,}
\end{split}
\eeq
where $R_g$ is the guiding radius of the orbit.

An induced error in $z$ of $\Delta z$ will cause $\Delta A_z \sim \Delta z$.
Similarly, an induced error in $R$ will cause $\Delta A_R \sim \Delta R$. If
we assume that $\kappa$ and $\nu$ are constant (\Gus{justify?}), then we have
that
\beq\label{eq:actions_epi_error}
\begin{split}
\Delta L_z = \Delta R v_c &\implies \frac{\Delta L_z}{L_z} = \frac{\Delta R}{R} \\
\Delta J_r = 2 \kappa \Delta A_R &\implies \frac{\Delta J_r}{J_r} = \frac{2\Delta R}{A_R} \\
\Delta J_z = 2 \nu \Delta A_z &\implies \frac{\Delta J_z}{J_z} = \frac{2\Delta z}{A_z}\text{.}
\end{split}
\eeq
Therefore, errors in each action are linearly proportional to the errors in
either the midplane or cylindrical radius.

>>>>

Under this approximation, we can write down the vertical and radial components
of the orbits as,
\beq\label{eq:orbits_epi}
\begin{split}
z(t) &= A_z \sin{(\kappa t + \alpha)}\\
R(t) &= A_R \sin{(\nu t + \beta)}\text{,}
\end{split}
\eeq
where $\kappa$ and $\nu$ are the vertical and radial (or epicyclic)
frequencies, $A_z$ and $A_R$ are the amplitudes of the stellar motion in the
vertical and radial coordinates, and $\alpha$ and $\beta$ are arbitrary
constants. Similarly, the velocities of the orbit are given by,
\beq\label{eq:orbits_vel_epi}
\begin{split}
v_z(t) &= A_z \kappa \cos{(\kappa t + \alpha)}\\
v_R(t) &= A_R \nu \cos{(\nu t + \beta)}\text{.}
\end{split}
\eeq

We can write down the actions in this approximation as
\citep[Section~3.5.3b][]{2008gady.book.....B},
\beq\label{eq:actions_epi}
\begin{split}
L_z &= R_g v_c \\
J_r &= \kappa A_R^2 \\
J_z &= \nu A_z^2\text{,}
\end{split}
\eeq
where $R_g$ is the guiding radius of the orbit.

An induced error in $z$ of $\Delta z$ will cause $\Delta A_z \sim \Delta z$.
Similarly, an induced error in $R$ will cause $\Delta A_R \sim \Delta R$. Making the assumption that $\kappa$ and $\nu$ are constant, we have
that
\beq\label{eq:actions_epi_error}
\begin{split}
\Delta L_z = \Delta R v_c &\implies \frac{\Delta L_z}{L_z} = \frac{\Delta R}{R} \\
\Delta J_r = 2 \kappa \Delta A_R &\implies \frac{\Delta J_r}{J_r} = \frac{2\Delta R}{A_R} \\
\Delta J_z = 2 \nu \Delta A_z &\implies \frac{\Delta J_z}{J_z} = \frac{2\Delta z}{A_z}\text{.}
\end{split}
\eeq
Therefore, errors in each action are linearly proportional to the errors in
either the midplane or cylindrical radius. While in detail $\kappa$ and $\nu$ are not constant, we will find that this order of magnitude estimate matches the numerical estimates quite well.
